---
title: "Initial overview of data"
author: "MIMKATZ"
date: "March 31, 2016"
output: html_document
---

Prezentacja 1
========================================================


##Introduction
We read our data into R.

```{r}
library(plyr)
library(ggplot2)
```

```{r}
#data <- read.csv("~/Pobrane/exp_oceny.txt", sep=";", dec=",")
data <- read.csv("~/jnp2/proj/exp_oceny.txt", sep=";", dec=",")
data <- data[, c(1:10)]
```

##Initial overview
Let's check the dimensions of our data.
```{r}
nrow(data)
ncol(data)
```

Most of them are empty, so we cut off the useless ones.

```{r}
data <- data[, c(1:10)]
```

Let's see what are the names of our columns.

```{r}
names(data)
```

Let's see the data type of our columns.

```{r}
sapply(data, class)
```

We select subjucts from our faculty, excluding OGUNs etc.

```{r}
data <- data[grepl("1000-*", data$PRZ_KOD),]
```

This is how our data looks

```{r}
head(data)
```

Let's check the count of each of the values

```{r}
count <- ddply(data, ~ PRZ_KOD + NAZWA, nrow)
count <- arrange(count, desc(V1))
head(count)
```

It appears that many people from MIM attend History of maths. That's not too surprising though.
There seems to be much data about analytic functions (FAN), let's see what's up there.

So we select only FAN-related data.
```{r}
FAN <- data[grepl("*FAN$", data$PRZ_KOD),]
```

Let's generate the simplest histogram with grades to get an overview.
```{r}
plot <- ggplot(data=FAN, aes(x=OCENA_WARTOSC)) + geom_bar(position="dodge")
```

Let's additionally divide it by academic semester.
```{r}
plot + facet_wrap(~ CDYD_KOD)
```

Thus we see that in 2008L and 2011Z something interesting is happening - in 2008L there are many 3 grades. Is that maybe related to a strict lecturer or to a weak year-group. On the other hand, in 2011Z there were many 5 grades - once again, is that a question of a skilled year-group or is it related to an easy-going lecturer.

Let's check it out from the POV of the lecturers/tutors.
```{r}
plot + facet_wrap(~ PRAC_ID_1)
```

Persons no 208 and 379 seem conspicuous, might they have been the lecturers/tutors in 2008L and 2011Z?
Let's limit ourselves to 2008L first.
```{r}
FAN_2008 <- FAN[ grepl("^2008L$", FAN$CDYD_KOD),]
table(FAN_2008$PRAC_ID_1)
```
Indeed, 208 tutored then.
Let's divide the grades by lecturers in 2008 - was it only one person that gave many 3s, or was the year-group bad?
```{r}
plot <- ggplot(data=FAN_2008, aes(x=OCENA_WARTOSC, fill=PRAC_ID_1)) + geom_bar(position="dodge")
plot + facet_wrap(~ PRAC_ID_1)
```

It seems to have been the second suspicioun. Or a hard exam.

Just to be on the safe side, let's check the distribution of the tutors during all semesters.
```{r}
FAN_teacher <- FAN[ grepl("^208$|^379$|^278$|^109$", FAN$PRAC_ID_1),]
```

```{r}
plot <- ggplot(data=FAN_teacher, aes(x=OCENA_WARTOSC, fill=PRAC_ID_1)) + geom_bar(position="dodge")
```

```{r}
plot + facet_wrap(~ CDYD_KOD)
```
In 2008 we have much more data than in other years, this definitely skews our perception. However, during the other years one doesn't see any significant deviations, only in 2008 there are many 3s.

```{r}
plot + facet_wrap(~ PRAC_ID_1)
```

Let's leave maths and concentrate on something interesting for us programists.
Let's start with ASD.
```{r}
ASD <- data[grepl("*ASD$", data$PRZ_KOD),]
```

Once again, we check the grades with regards to dydactic cycle.

```{r}
plot <- ggplot(data=ASD, aes(x=OCENA_WARTOSC)) + geom_bar(position="dodge")
```

```{r}
plot + facet_wrap(~ CDYD_KOD)
```

The data for older years is rather incomplete - it's interesting to note that this subject used to be taught also in the summer semester and only in 2010 was it permanently moved to the winter semester. Let's concentrate our attention on the 3 year-groups that we have the most information on.

```{r}
ASD <- ASD[ grepl("^2010Z$|^2011Z$|^2012Z$", ASD$CDYD_KOD),]
```

```{r}
plot <- ggplot(data=ASD, aes(x=OCENA_WARTOSC)) + geom_bar(position="dodge")
```

The number of very good grades among the years is similar, however the number of fail grades in 2011 is twice as many as in the year before.

```{r}
plot + facet_wrap(~ CDYD_KOD)

```

Let's do the same for operating systems (SO).

```{r}
SO <- data[grepl("*SOP$", data$PRZ_KOD),]
plot <- ggplot(data=SO, aes(x=OCENA_WARTOSC)) + geom_bar(position="dodge")
plot + facet_wrap(~ CDYD_KOD)
```


```{r}
SO <- SO[ grepl("^2010Z$|^2011Z$|^2012Z$", SO$CDYD_KOD),]
plot <- ggplot(data=SO, aes(x=OCENA_WARTOSC)) + geom_bar(position="dodge")
plot + facet_wrap(~ CDYD_KOD)
```

At first glance, it seems similar as with ASD, but maybe not?

```{r}
ASD_SO <- rbind(ASD, SO)
plot <- ggplot(data=ASD_SO, aes(x=OCENA_WARTOSC, fill=NAZWA)) + geom_bar(position="dodge")
plot + facet_wrap(~ CDYD_KOD)

```

For a change, let's lok at the year 2010 and these three subjects: MD, ASD and SIK.

```{r}
ALL_2010 <- data[data$CDYD_KOD == c("2010Z", "2010L"),];
INF_ob <- ALL_2010[grepl("*MAD$|*SIK$|*ASD$", ALL_2010$PRZ_KOD), ];
```

Let's compare the grades across these subjects.

```{r}
plot <- ggplot(INF_ob, aes(x = NAZWA, y = OCENA_WARTOSC)) + geom_dotplot(binaxis = "y", stackdir = "center")
plot
```

Let's see if there's some link between grades in WPI and ASD. Let's select people that have grades in both (only two people).

```{r}
inf <- data[grepl("*211aWPI$|*aASD$", data$PRZ_KOD), ]

o <- ddply(inf, ~ OS_ID, nrow)
o <- arrange(o, desc(V1))
o <- o[o$V1==2,]

inf <- inf[inf$OS_ID %in% o$OS_ID, ]

```

And a plot.

```{r}
plot <- ggplot(data=inf, aes(x=NAZWA, y=OCENA_WARTOSC, group=OS_ID))
plot + geom_line(size=2) + geom_point(size=4)
```

The same for SO.

```{r}
inf <- data[grepl("*211aWPI$|*215aSOP$", data$PRZ_KOD), ]

o <- ddply(inf, ~ OS_ID, nrow)
o <- arrange(o, desc(V1))
o <- o[o$V1==2,]

inf <- inf[inf$OS_ID %in% o$OS_ID, ]
```


```{r}
plot <- ggplot(data=inf, aes(x=NAZWA, y =OCENA_WARTOSC, group=OS_ID))
plot + geom_line(size=2) + geom_point(size=4)
```

As we can see on the previous plots, the data wasn't too good. Indeed, there were two persons with two grades, but both were from the same subject. This mistake englightened us to make two new plots that would answer the question: how do people with more that one grade in a subject fare in subsequent years.

First we check this out for ASD.

```{r}
inf <- data[grepl("*aASD$", data$PRZ_KOD), ]

o <- ddply(inf, ~ OS_ID, nrow)
o <- arrange(o, desc(V1))
o <- o[o$V1>1,]

inf <- inf[inf$OS_ID %in% o$OS_ID, ]
```


```{r}
plot <- ggplot(data=inf, aes(x=CDYD_KOD, y=OCENA_WARTOSC, group=OS_ID, color=OS_ID))
plot + geom_line(size=2) + geom_point(size=4)
```

And next for SO.

```{r}
inf <- data[grepl("*215aSOP$", data$PRZ_KOD), ]

o <- ddply(inf, ~ OS_ID, nrow)
o <- arrange(o, desc(V1))
o <- o[o$V1>1,]

inf <- inf[inf$OS_ID %in% o$OS_ID, ]

```


```{r}
plot <- ggplot(data=inf, aes(x=CDYD_KOD, y=OCENA_WARTOSC, group=OS_ID, color=OS_ID))
plot + geom_line(size=2) + geom_point(size=4)
```